On the eletrodynamic of moving bodies

### 2019-06-12 Article_AS.Rmd
Intro apres commentaires generaux de AA

---
title: "On the eletrodynamic of moving bodies"
author: "André Dogninema Soro, Jean-Romain Roussel, Alexis Achim"
output: 
  pdf_document: default
  word_document: default
bibliography: biblio.bibtex
editor_options: 
  chunk_output_type: inline
---

@@

Canadian forest is an important part of the country's history, culture, natural environment and economy. But thousands of hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. To maintain resource constant, it is therefore essential to compensate the harvest and the looses by reforesting. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use. \

@@

During the first half of the 20th century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Wood quality has a biologic and an economic component. Biologically, the resistance to mortality by vigour is important to ensure to harvest leaving trees. Economically the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the homogeneity of the density within the stem has an high importance. Its has been demonstrated that trees with a more uniform wood density within the stem are also the more resistant to mortality [@Hietz2013; @osazuwa2014] because a uniform density gives a greater mechanical resistance to the wind and therefore a lower probability of mortality during a climatic event such as a storm. With regard to the economical context, variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. In fact, wood density uniformity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bio energy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, a wide inter-annual variation of the density of the wood leads to a decrease of the mechanical resistance while a relatively uniform density produce wood with better mechanical performance [@x,@y]. It is therefore important from both a biological and an end-use points of view to have a wood with a uniform density within the stem. \

In view of the studies cited above, we understand that there might be a strong gain at selecting trees that are prone to produce wood with an homogeneous density to get a resistant and highly valuable forest in the next decades. Consequently, studies have been set up to understand the genetic sources of phenotype variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Mullin2011; @Beaulieu2009]. Hence, it is important to understand where do the variations of density come from in order to better integrated them into genetic selections. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011] with a density that decreases during the first years of growth and then an increasing from the beginning of the formation of mature wood [@panshin1980]. These variations are under genetic control and are consequently moderately to highly heritable [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. This makes this trait attractive for genetic improvement. In addition to these low-frequency variations, there are high frequency inter-annual variations [@Nicault2010] that are influenced mainly by events to the tree such as the edaphic and climatic conditions [@Nicault2010] that happen at high frequency. These variations are not under genetic control but under climate control meaning that climatic conditions also affect wood density. Some climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. Long-term growth analyses have shown that density may even increases with the temperature, the atmospheric CO~2~ content and the nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Wood density can also be negatively correlated with growth. [@zhang1995]. Therefore, with the aim of producing wood of homogeneous density it is essential to take climate factors into account in addition to genetic.

The impact of climate is more perceptible on the properties of wood when there are wide variations of it such as drought, high temperatures or significant rainfall. The ability of trees to tolerate these changes in a tree species or within a species is often related to the interaction between genotype and environment. For example, studies have shown a low sensibility to genotype-related drought [@ferrell1966; @teskey1987]. Consequently, the sensitivity of trees to climate such as inter-annual variability in density can be controlled by genetics. It is therefore important to use trees that are less sensitive to climate variations because of their genome for breeding programs.

@@

## Sampling
The dataset used in this study came from genetic selection for a series of tests according to provenances [@Beaulieu1996]. The improvement program for this region was led by Quebec’s Ministry of Forests, Wildlife and Parcs, Canada. Experimental plantings were initiated from white spruce seedlings grown in nurseries and planted at two years of age at a regular spacing between trees of 2 m in all directions. This series of tests was established in 1999 by the Canadian Forest Service and is one of the three essential parts of the white spruce breeding program in Quebec. Two sites were selected for sampling: one near Saint-Casimir, Quebec (46.6994, -72.1119), in the maple–basswood ecological zone (warm weather) , and the other one near Cabano (Asselin), Quebec (47.8808, -68.4436), in the balsam fir-yellow birch ecological zone (cool weather) (fig 1). Between 9 and 14 trees per site were sampled from each of the 93 full-sib families, for a total sample of 2365 trees. Tree height and diameter at breast height (DBH, 1.3 m above ground) were recorded in 2015 at the age of 15 years, and a wood increment core was extracted from the south facing side of each tree. Cores were stored in a freezer, conditioned to 7% moisture and cut to 1.68 mm thickness prior to X-ray densitometry analyses (Quintek Measurement Systems, TN) at a resolution of 0.0254 mm, and the data were reported as relative wood density on a dry weight basis. Wood density was calculated as a ring area weighted mean from recorded pith to bark wood density profiles.

@@

To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for cross-dating and quality control of annual ring measurements [@HOLMES1983]. thus, as the annual ring separation was carried out by an algorithm, we eliminated the samples (127 samples) whose dating was not correct and could not be modified.

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce. Considering the hierarchical structure of the data, with annual rings nested in the cores, cores nested in the trees, and trees nested in blocks, we opted for a mixed-effects modelling approach using nlme package [@NLME2018] in the R statistical programming environment [@R2018]. This type of model also considers the autocorrelation between the observations [@Lindstrom1990]. Thus, the estimates of the parameters can vary randomly around the average values of each level of the data structure (site, block, tree, core) [@Auty2013]. As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

@@

## Spearman’s rank correlation coefficient (Spearman’s rho) 
To describe the link between two variables by means of a monotonic function, we used Spearman’s rank correlation coefficient. We calculated this coefficient for the index of variability and the different parameters of the model. The elements for which we calculated Spearman’s rho are: the variability index for each of the tree families between two study sites, the model residues for each of the families and the three parameters to estimate from the model. The cor.test function in the R [@R2018] statistical programming environment was used to calculate Spearman’s rho. We used a 95 percent confidence interval.

## Quantitative genetic analyses
Variance and covariance for genetics were estimated using mixed models [@Lenz2013] of the R software. The model used is as follows:

@@

where $yijkl$  is the observation on the $ijklth$ tree, $\mu$ is the overall mean, $s_i$ is the fixed site effect, $b_{(i)j}$ is the fixed effect of the $j^{th}$ bloc in site $i$, $f_k$ is the random effect of the $k^{th}$ family, $bf_{(i)j}$ is the random interaction of family and bloc, and $e_{ijkl}$ is the residual error.
In a full-sib design with random mating, members of the same family are presumed to share half of their genes [@white2007]. Therefore, individual heritability was obtained as follows:

 $$h_2= 2 \frac{V_f}{V_T}$$

Where $h_2$ is individual heritability, $V_f$ is family variance, $V_T$ is total variance.
ASReml3 package [@Butler2018] in the R statistical programming environment [@R2018] was used to estimate trait genetic parameters based on the residual maximum likelihood.

@@

The model we have developed has made it possible to optimize the density of the wood for each study site. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. For the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin.\

@@

When considering all the cores of trees from the pith-to-bark, there is initially a rapid decrease in density in the first five years. It goes from 430 kg/m^3^ to 370 kg/m^3^ for the Saint-Casimir site and from 455kg/m^3^ to kg/m^3^ for the Asselin site. Then, this decrease was followed by a slight increase until the twelfth year.\

@@

|Trait            |Combined sites |Asselin |Saint-Casimir |
|:----------------|:--------------|:-------|:-------------|
|Wood density     |0.55           | 0.58   |0.70          |
|Width            |0.23           |0.32    |0.43          |
|$b_0$            |0.20           |0.20    |0.3           |
|$b_1$            |0.06           |0.10    |0.06          |
|$b_2$            |0.04           |0.10    |0.03          |
|Uniformity index |1.10^-6^       |6.10^-6^|3.10^-6^      |
Table: Abstract of the heritability of the density, the ring width, parameters $b_0$, $b_1$, $b_2$ and the Variability index.

@@

The data set showed a typical radial pattern of wood density already observed in spruce trees in other studies [@Alteyrac2006; @GARDINER2011; @Xiang2014]. The trend is similar to that described as Type II density profile in @panshin1980. The high density during the first years of tree growth can be explained by the immature tracheids combined with the high proportion of compression wood. The density decreases to a certain level and begins to increase again. This increase is due to the fact that the width of the ring decreases from a certain age, so the proportion of initial decreases and that of the final wood [@Lachenbruch2011]. 
The modeling of individual profiles is suitable for the description of the wood density of our type II data. The choice of model allowed to take into account the variability of the density profiles of white spruce. @Xiang2014 applied a similar model in their study to describe the density profiles of black spruce with density profiles similar to those of white spruce. The low index of parameters $b_1$ and $b_2$ were most likely the large variability of the initial decline near the pith. Although the variaton of density follows a Type II profile, it remains slightly different between the two sites. This difference can be explained by the difference in pedoclimatic cinditions between sites. Although, the soil type is different between sites and the climate is colder in Asselin compared to Saint-Casimir. At the level of the model parameters, the sites are also different and this difference can be explained by the profiles of densities that are not the same.\

>>>>

---
title: "On the eletrodynamic of moving bodies"
author: "André D. Soro, Patrick Lenz, Jean Bousquet, Mariana Hassegawa, Jean-Romain Roussel, Alexis Achim"
output: 
  pdf_document: default
  word_document: default
bibliography: biblio.bibtex
editor_options: 
  chunk_output_type: inline
---

@@

Thousands hectares of Canadian forest are harvested each year and an additional loss is due to insects, diseases and fires [@DRF2000; @NaturalResourcesCanada2019]. It is important to increase tree growth to meet the growing demand of forest industry. The reforestation, when under human control, could ideally be made by using better trees than the former ones, i.e. trees that growth faster, that are more resistant to perturbation and that are of better structural quality. To achieve this task, we must take into account the cambial activity that is governed by environmental conditions and genetics [@Dufour2010; @Rossi2012]. Therefore, genetic selection is one of the available tools and is currently becoming a crucial and increasingly necessary for the development of the canadian forest [@Bonfils2004]. It allows to select trees with characters of interest and to cross them or to reproduce them in order to obtain trees with interesting properties for the development of the forests or the quality of the wood for the human use [@zobel1984; @DRF2000; @nanson2004]. \

@@

During the first half of the 20^th^ century, most studies about genetic selection were dedicated on increasing wood volume production [@Dhir1976; @nienstaedt1985; @Li1997; @Koubaa2000] but with the increasing demand of high structural quality products and the fear of losing resource from natural causes, the aspects of wood quality became as important as wood volume [@Franceschini2012]. Therefore, conifer breeding programs are more and more considering wood quality as selection criteria to improve end-use properties. Stem vigor (biologic component) and wood quality (economic component) are two desirable traits that are taken into account in genetic selection. Biologically, the resistance to mortality by vigour is important to ensure to harvest leaving trees. Economically the value of the wood for end-use is important for industries. Several traits were considered to assess wood quality and the density is one of the most widely used indicators and arguably the most important one [@zobel1989; @Saranpaa1994; @Mckendry2002; @Macdonald2002; @Alteyrac2006]. More specifically the homogeneity of density and rings width within the stem has an high importance for stem vigor and wood quality because these two traits are related to it. Homogeneity refers here to a small interannual variation in wood density and rings width.\

In view of the studies cited above, we understand that there might be a strong gain at selecting trees that are prone to produce wood with an homogeneous density to get a resistant and highly valuable forest in the next decades. Consequently, studies have been set up to understand the genetic sources of phenotype variations of wood quality [@Corriveau1971; @Li1993; @Hernandez2001; @Beaulieu2003; @Lenz2013; @McLean2016]. For example, future genetic selections in breeding programs for different species of spruce trees in Quebec will consider wood properties such as wood density [@Beaulieu2009; @Mullin2011]. Hence, it is important to understand where do the variations of density come from in order to better integrated them into genetic selections. In white spruce, there is a typical low frequency radial pattern of variation from pith to bark [@Moore2011] with a density that decreases during the first years of growth and then an increasing from the beginning of the formation of mature wood [@panshin1980]. These variations are under genetic control and are consequently moderately to highly heritable [@Corriveau1991; @Yanchuk1993; @Ivkovich2002a; @Ivkovich2002b; @Beaulieu2006]. This makes this trait attractive for genetic improvement. In addition to these low-frequency variations, there are high frequency inter-annual variations [@Nicault2010] that are influenced mainly by events to the tree such as the soil water availability and climatic conditions [@Nicault2010] that happen at high frequency. These variations are not under genetic control but under climate control meaning that climatic conditions also affect wood density. Some climatic factors such as drought can cause stress in trees and cause a variation in wood properties, including density, ring width and conductivity. Long-term growth analyses have shown that density may even increases with the temperature, the atmospheric CO~2~ content and the nitrogen deposition [@becker1995; @huang2007; @zaehle2014]. Wood density can also be negatively correlated with growth. [@zhang1995]. Therefore, with the aim of producing wood of homogeneous density it is essential to take climate factors into account in addition to genetic.\

 In fact, the density is correlated with the growth of the tree [@corriveau1991] and it has been observed that trees with a small variation in growth and therefore in density are also the most resistant to mortality [@ogle2000; @suarez2004; @hartmann2008]. For example, @suarez2004 found that at the tree scale, individuals with variable growth were more prone to die of drought than more regularly growing trees. These studies have shown that the steady growth and therefore the small variation in the density of the wood allows the tree to be more vigorous. With regard to the economical context, variation of density within a piece of wood may lead to deformation during transformation and inhomogeneous machining properties. Indeed, wood density homogeneity affects the performance of saw timber, the pulp yield and calorific values for biomass applications for bio energy [@Hernandez2001; @Macdonald2002; @Mckendry2002]. For example, the low variation in wood density is appreciated for sawnwood because sawyers place a high value on the homogeneity of wood density [Hassegawa et al (pending approval); @X]. It is therefore important from both a biological and an end-use points of view to have a wood with a uniform density within the stem. \

A significant variation of the climate can cause a variation of the properties of the wood over several years. For example, several studies have shown that after the great drought of 2003 in Europe, the growth of forests has been affected over several years [@ciais2005; @granier2007; @reichstein2007]. The ability of trees to tolerate these changes in a tree species or within a species is often related to the interaction between genotype and environment. For example, studies have suggested a possible link between low sensitivity to drought and genotype [@ferrell1966; @teskey1987]. This link is poorly controlled and therefore poorly known. It is in context that this study fits to try to advance the knowledge in this direction.\

@@

## Sampling
The dataset used in this study came from genetic selection for a series of tests according to provenances [@Beaulieu1996]. The improvement program for this region was led by Quebec’s Ministry of Forests, Wildlife and Parcs, Canada. Experimental plantings were initiated from white spruce seedlings grown in nurseries and planted at two years of age at a regular spacing between trees of 2 m in all directions. This series of tests was established in 1999 by the Canadian Forest Service and is one of the three essential parts of the white spruce breeding program in Quebec. Two sites were selected for sampling: one near Porneuf (Saint-Casimir), Quebec (46.6994, -72.1119), in the maple–basswood ecological zone (warm weather) , and the other one near Cabano (Asselin), Quebec (47.8808, -68.4436), in the balsam fir-yellow birch ecological zone (cool weather) (fig 1). Between 9 and 14 trees per site were sampled from each of the 93 full-sib families, for a total sample of 2365 trees. Tree height and diameter at breast height (DBH, 1.3 m above ground) were recorded in 2015 at the age of 15 years, and a wood increment core was extracted from the south facing side of each tree. Cores were stored in a freezer, conditioned to 7% moisture and cut to 1.68 mm thickness prior to X-ray densitometry analyses (Quintek Measurement Systems, TN) at a resolution of 0.0254 mm, and the data were reported as relative wood density on a dry weight basis. Wood density was calculated as a ring area weighted mean from recorded pith to bark wood density profiles.

@@

To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for cross-dating and quality control of annual ring measurements [@HOLMES1983]. Thus, as the annual ring separation was carried out by an algorithm, we eliminated the samples (127 samples) whose dating was not correct and could not be modified. Any incomplete or false rings were excluded from the dataset.

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce. Considering the hierarchical structure of the data, with observations on annual rings nested in the cores, cores nested in the trees, and trees nested in blocks, we opted for a mixed-effects modelling approach using nlme package [@NLME2018] in the R statistical programming environment [@R2018]. This type of model also considers the autocorrelation between the observations [@Lindstrom1990]. Thus, the estimates of the parameters can vary randomly around the average values of each level of the data structure (site, block, tree, core) [@Auty2013]. As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

@@

## Spearman’s rank correlation coefficient (Spearman’s rho) 
To describe the link between two variables by means of a monotonic function, we used Spearman’s rank correlation coefficient. We calculated this coefficient for the variability index and the different parameters of the model. The elements for which we calculated Spearman’s rho are: the variability index for each of the tree families between two study sites and the three parameters to estimate from the model. The $cor.test$ function in the R [@R2018] statistical programming environment was used to calculate Spearman’s rho. We used a 95 percent confidence interval.

## Quantitative genetic analyses
Variance and covariance for genetics were estimated using mixed models [@Lenz2013] of the R software. The equation used is as follows:

@@

where $y_{ijkl}$  is the observation on the $ijklth$ tree, $\mu$ is the overall mean, $s_i$ is the fixed site effect, $b_{(i)j}$ is the fixed effect of the $j^{th}$ bloc in site $i$, $f_k$ is the random effect of the $k^{th}$ family, $bf_{(i)j}$ is the random interaction of family and bloc, and $e_{ijkl}$ is the residual error.
In a full-sib design with random mating, members of the same family are presumed to share half of their genes [@white2007]. Therefore, individual heritability was obtained as follows:

 $$h_2= 2 \frac{\sigma^2_f}{\sigma^2_T}$$

Where $h_2$ is individual heritability, $\sigma^2_f$ is family variance, $\sigma^2_T$ is total variance.
ASReml3 package [@Butler2018] in the R statistical programming environment [@R2018] was used to estimate trait genetic parameters based on the residual maximum likelihood.

@@

The model we have developed has made it possible to optimize the density of the wood for each core. In table 1, we observe that the average value of the parameters differs between Asselin and Saint-Casimir. For the parameters $b_0$ and $b_1$, the average is higher in Saint-Casimir and for parameter $b_2$ it is higher in Asselin.\

@@

When considering all the cores of trees from the pith-to-bark, ring density was high near the pith and decreased rapidly in the first 5 growth rings. It goes from 430 kg/m^3^ to 370 kg/m^3^ for the Saint-Casimir site and from 455kg/m^3^ to kg/m^3^ for the Asselin site. Then, this decrease was followed by a slight increase until the twelfth year.\

@@

|Trait            |Combined sites |Asselin |Saint-Casimir |
|:----------------|:--------------|:-------|:-------------|
|Wood density     |0.55           | 0.58   |0.70          |
|Width            |0.23           |0.32    |0.43          |
|$b_0$            |0.20           |0.20    |0.3           |
|$b_1$            |0.06           |0.10    |0.06          |
|$b_2$            |0.04           |0.10    |0.03          |
|Variability index |1.10^-6^       |6.10^-6^|3.10^-6^      |
Table: Abstract of the heritability of the density, the ring width, parameters $b_0$, $b_1$, $b_2$ and the Variability index.

@@

The data set showed a typical radial pattern of wood density already observed in spruce trees in other studies [@Alteyrac2006; @GARDINER2011; @Xiang2014]. The trend is similar to that described as Type II density profile in @panshin1980. The high density during the first years of tree growth can be explained by the immature tracheids combined with the high proportion of compression wood. The density decreases to a certain level and begins to increase again. This rapid decline in density is due to the significant growth of annual rings and the following increase is explained by the fact that the ring width  decreases from a certain age, because of the proportion of earlywood and latewood by ring.[@Lachenbruch2011].\

The modeling of individual profiles is suitable for the description of the wood density of our type II data. The choice of model allowed to take into account the variability of the density profiles of white spruce. @Xiang2014 applied a similar model in their study to describe the density profiles of black spruce with density profiles similar to those of white spruce. \

Although the variaton of density follows a Type II profile, it remains slightly different between the two sites. This difference can be explained by the difference in pedoclimatic conditions between sites. In fact, the soil type is different between sites and the climate is colder in Asselin which is further south compared to Saint-Casimir. At the level of the model parameters, the sites are also different and this difference can be explained by the profiles of densities that are not the same.\
